---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Lets load some libraries that we'll later.

```{r}
library(dplyr)
```

If activity.csv doesn't exist unzip it from the ZIP file.  

```{r}
if (!file.exists("activity.csv")) {
        unzip("activity.zip")
}
 
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

Use the aggregate function to find the total number of steps taken per day.  

```{r}
steps <- aggregate(activity$steps, by=list(Date=activity$date), FUN=sum)
```

The instructions said to produce a "histogram", but it sounded like what they really want is a barplot.  So I provided both.  

Histogram:
```{r}
hist(steps$x)
```

Bar Plot:
```{r}
barplot(steps$x, main = "Steps per Day", xlab = "Date", ylab = "Steps")
```

The mean and median steps per day are:  

```{r}
mean(steps$x, na.rm = TRUE)
median(steps$x, na.rm = TRUE)
```


## What is the average daily activity pattern?

Plot of the average steps per interval:

```{r}
stepsbyinterval <- aggregate(activity$steps, by=list(interval=activity$interval), FUN=mean, na.rm = TRUE)

intervals <- strptime(sprintf("%04d",as.numeric(stepsbyinterval$interval)), format = "%H%M")

plot(intervals, stepsbyinterval$x, type = "l", xlab="Interval (hh:mm)", ylab="Steps per Interval")
```  

The interval with the maximum number of steps per day, on average is:

```{r}

stepsbyinterval[ stepsbyinterval$x == max(stepsbyinterval$x), ]

```


## Imputing missing values

To impute missing data I plan to use the average for a given interval.  

First compute the mean steps for each interval:

```{r}
aveSteps <- aggregate(activity$steps, by=list(Date=activity$interval), FUN=mean, na.rm=TRUE)
```

Give the column meaningful names.

```{r}
names(aveSteps) <- c("interval", "mean.steps")
```

Lets create a copy of the activity dataframe and add to it a column that contains the 
average number of steps for that interval.

```{r}
act <- merge(activity, aveSteps, by = "interval")
head(act)
```

Create a new dataset using steps where it's available and mean.steps where steps is NA.

I've created a small function to help me select where the steps value is NA.

```{r}
stepfunc <- function(a,b){
        
        if( is.na(a) ) {
                b
        } else {
                a
        }
        
} 
```

Force creation of a vector "a".
```{r}
a<-0
```

Now populate "a" with either the number of steps or, if teh number of steps is NA, the mean number of steps for that interval.
```{r}
for(i in  1:nrow(act)) {
        a[i]<-stepfunc(act[i,2], act[i,4])
        
}
head(a, n = 20L)
```

Add the new column to the "act" dataset.
```{r}
act<- cbind(act,a)
head(act, n = 20L)
```

## Are there differences in activity patterns between weekdays and weekends?




